---
title: "DS6306 Case Study 01 - Beers and Breweries"
authors: "Lauren Nelson, Alan Abadzic, Robert Derner"
date: "10/9/21"
output: html_document
---

```{r setup}
knitr::opts_chunk$set(echo = TRUE)
library(usmap)
library(ggplot2)
library(maps)
library(ggsn)
library(plotly)
library(tidyverse)
```

# Introduction
This provides insight and analysis into Beer and Brewery data across the US.  Data was provided by Budweiser Executives for strategic decision making.

```{r}
#library(tidyverse)
beers = read_csv(url("https://raw.githubusercontent.com/BivinSadler/MSDS_6306_Doing-Data-Science/Master/Unit%208%20and%209%20Case%20Study%201/Beers.csv"))
breweries = read_csv(url("https://raw.githubusercontent.com/BivinSadler/MSDS_6306_Doing-Data-Science/Master/Unit%208%20and%209%20Case%20Study%201/Breweries.csv"))

summary(beers)
summary(breweries)

head(beers)
head(breweries)
```

1.	How many breweries are present in each state?
# Breweries by State
```{r}

#state_data -- Breweries Filtered by State - joined with FIPS code by State
#mutate - Join FIPS data with Breweries data
#trimws - remove Leading/Trailing Whitespace from column(State)
#fips = State by Abbreviation - ie. TX

state_data = breweries %>% count(State) %>% mutate(fips = fips(trimws(State)))

us_map <- usmap::us_map() # used to add map scale

usmap::plot_usmap(data = state_data, values = "n", labels = T)+
  labs(fill = 'Breweries by State') + 
  scale_fill_gradientn(colours=rev(topo.colors(50)),na.value="grey90",
                       guide = guide_colourbar(barwidth = 25, barheight = 0.6,
#put legend title on top of legend
title.position = "top")) +

  # put legend at the bottom, adjust legend title and text font sizes
  theme(legend.position = "bottom",
        legend.title=element_text(size=12), 
        legend.text=element_text(size=10))
```


2.	Merge beer data with the breweries data. Print the first 6 observations and the last six observations to check the merged file.  (RMD only, this does not need to be included in the presentation or the deck.)


```{r}
#Merge beer and breweries data - trim whitespace - add in FIPS codes
full_data = merge(beers,breweries,by.y="Brew_ID",by.x="Brewery_id") %>%
  mutate(fips = fips(trimws(State)))

head(full_data)
tail(full_data)
full_data[sample(1:nrow(full_data),6),]
```

3.	Address the missing values in each column.

Two Datasets, "Beers.csv" and "Breweries.csv" were joined to create a Full Dataset for analysis and insight

Only ABV and IBU are missing Data Values

Important to note: IBU is missing values in  around 42% of the 2,410 total rows of data

**Predictions for IBU specifically, may be skewed

```{r}
#use colSum to sum up the NA's in each column
full_dataNAs <- as.data.frame(colSums(is.na(full_data)))
names(full_dataNAs) <- c("Missing Values (NAs)")
full_dataNAs

summary(full_data)

1005/2410 #Calculate percentage missing data from IBU Column
```

4.	Compute the median alcohol content and international bitterness unit for each state. Plot a bar chart to compare.
# Alcohol Content, Bitterness by State - BarChart

```{r}
#Preliminary Steps

#Import Beer and Brewery Data Sets
Beer = read.csv("https://raw.githubusercontent.com/BivinSadler/MSDS_6306_Doing-Data-Science/Master/Unit%208%20and%209%20Case%20Study%201/Beers.csv", header = TRUE)
Breweries = read.csv("https://raw.githubusercontent.com/BivinSadler/MSDS_6306_Doing-Data-Science/Master/Unit%208%20and%209%20Case%20Study%201/Breweries.csv", header = TRUE)

#Updating the name of the first column of the breweries dataset so there is a matching column between the two datasets
names(Breweries)[1] <- 'Brewery_id'

#Merging the two sets together by the Brewery ID column
BB = merge(Beer,Breweries, by="Brewery_id", all = TRUE)
#Updating the mame of the 8th column so that it is Brewery
names(BB)[8] <- 'Brewery'
#Omitting all N/A values from the ABV and IBU columns
x = na.omit(BB[,c(4,10)])
y = na.omit(BB[,c(5,10)])
```

Alcohol by Volume Comparison by State

```{r}
#Finding the median ABV of each state
medians = aggregate(x=x$ABV, by = list(x$State), FUN = median)
#Finding the median IBU of each state
mediansIBU = aggregate(x=y$IBU, by = list(y$State), FUN = median)
#Updating column names of the median ABV dataset
names(medians)[1]<- 'State'
names(medians)[2]<- 'ABV'
#Updating column names of the median IBU dataset
names(mediansIBU)[1]<- 'State'
names(mediansIBU)[2]<- 'IBU'

#Plotting the median ABV of each state
p = medians %>%
  arrange(desc(ABV, State)) %>%
ggplot(mapping = aes(x=State, y=ABV)) + geom_bar(stat = 'identity', fill = topo.colors(51), width = .5,color = 'black', show.legend = TRUE) + labs(y= "Median ABV", title = "Median ABV (Beers) by State") 

ggplotly(p) #Add interactive data labels
```
Bitterness Comparison by State 

```{r}
#Plotting the median IBU of each state
p = mediansIBU %>%
    arrange(desc(IBU, State)) %>%
ggplot(mapping = aes(x=State, y=IBU)) + geom_bar(stat = 'identity', fill = topo.colors(50), width = .5,color = 'black') + labs(y= "Median IBU", title = "Median IBU of Beers by State") 

ggplotly(p) #Add interacitve data labels
```

5a.	Which state has the maximum alcoholic (ABV) beer? 
#Max ABV = CO
```{r}
#Finding the highest ABV of all the states
MaxABV = aggregate(x$ABV, by = list(x$State),max)
#Finding the highest IBU of all the states
MaxIBU = aggregate(y$IBU, by = list(y$State),max)
#Updating column names of the max IBU dataset
names(MaxIBU)[1]<- 'State'
names(MaxIBU)[2]<- 'IBU'
#Updating column names of the max ABV dataset
names(MaxABV)[1]<- 'State'
names(MaxABV)[2]<- 'ABV'
#Order the MAX ABV and IBU datasets in descending order
MaxABV = BB[order(BB$ABV, decreasing = TRUE),]
MaxIBU = BB[order(BB$IBU, decreasing = TRUE),]

#Plotting the Max ABV of each state
p = MaxABV %>%
ggplot(mapping = aes(x=State, y=ABV)) + geom_bar(stat = "identity", fill = "darkslateblue") + labs(y="ABV", title = "Alcohol by Volume per State") 

ggplotly(p)
```

5b. Which state has the most bitter (IBU) beer?
#Most Bitter Beer = OR
```{r}
#Plotting the Max IBU of each state
p = MaxIBU %>%
ggplot(mapping = aes(x=State, y=IBU)) + geom_bar(stat = "identity", fill = "chartreuse3", width = .5,color = 'darkslateblue') + labs(y="IBU", title = "Highest Bitterness per State") 
ggplotly(p)
```

6.	Comment on the summary statistics and distribution of the ABV variable.

```{r}
#Budweiser

Beers = read_csv(url("https://raw.githubusercontent.com/BivinSadler/MSDS_6306_Doing-Data-Science/Master/Unit%208%20and%209%20Case%20Study%201/Beers.csv"))
Breweries = read_csv(url("https://raw.githubusercontent.com/BivinSadler/MSDS_6306_Doing-Data-Science/Master/Unit%208%20and%209%20Case%20Study%201/Breweries.csv"))





#merge the data before performing EDA
names(Breweries)[1] <- 'Brewery_id'
Breweries
Beer_Merge <- merge(x=Beers,y=Breweries, by = "Brewery_id", all.x = TRUE)

lookup = data.frame(abb = state.abb, State = state.name) #makes a data frame with State name and abbreviation.

summary(Beer_Merge$ABV) #see summary statistics of mean, median, mode, and quartiles
#plot the distribution of ABV
Beer_Merge %>%
  ggplot(aes(x = ABV)) + geom_density(binwidth = .5, colour="black",fill="chartreuse3") +
  geom_vline(aes(xintercept=mean(ABV, na.rm=T)),
             color = "red", linetype = "dashed", size = 1) +
  ggtitle("Distribution Density of ABV")
```

7.	Is there an apparent relationship between the bitterness of the beer and its alcoholic content? Draw a scatter plot.  Make your best judgment of a relationship and EXPLAIN your answer.

**There appears to a positive correlation between ABV and IBU however, IBU values may be skewed due to missing IBU data - 42% of total missing**

```{r}
Beer_Merge %>%
  ggplot(aes(x = ABV, y = IBU, fill = "sample")) +
  geom_point() + geom_smooth(method =  lm) +
  ggtitle("Relationship Between IBU and ABV")

correlation_test = cor.test(Beer_Merge$ABV,Beer_Merge$IBU) #run correlation to determine p value and r-sqaured
correlation_test #showcase statistics

#The scatterplot shows a positive correlation between level of bitterness and alcohol by volume indicating that the higher the bitterness the more alcohol that tends to be contatined within

#Even with a sample of 1005 beers out of 2410 observations in total of IBU, we can conclude that a p value of 2.2e-16 we can say,
#we should reject the null hypothesis given an intercept of 0, which is further proved by the confidence interval
#being .64 to .69 at alpha of .05 or 95%.
#There appears to be a strong relationship with a r-squared correlation of .6706 squared = .45.
#It's a fair to state that when the ABV or IBU is increased the other shall also increase.
```

##Bonus Insights

Top Colorado Beer Styles

```{r}
#Top 2 States _____#CO | #CA - Top Beer Styles for Top 2 States 

COStyle = full_data %>% filter(fips == "08") %>% select(State,Style) %>% group_by(State) %>% 
  count(Style) %>% slice_max(n,n=5) %>% mutate(fips = fips(trimws(State)))

#Colorado Top Beer Styles BarChart
p = COStyle %>% 
  arrange(desc(COStyle$Style)) %>% 
  mutate(Style = factor(Style)) %>% 
  ggplot(aes(y=Style,x=COStyle$n, fill = COStyle$n)) + geom_bar(stat='identity', show.legend = TRUE, width = .8) +
  labs(fill = 'CO Beer Styles') +
  ggtitle("Top Colorado Beer Styles")

ggplotly(p)
```

Top California Beer Styles

```{r}
CAStyle = full_data %>% filter(fips == "06") %>% select(State,Style) %>% group_by(State) %>% 
  count(Style) %>% slice_max(n,n=5) %>% mutate(fips = fips(trimws(State)))


#California Top Beer Styles BarChart
p = CAStyle %>% 
  arrange(desc(CAStyle$Style)) %>% 
  mutate(Style = factor(Style)) %>% 
  ggplot(mapping = aes(y=Style,x=CAStyle$n, fill = CAStyle$n)) + geom_bar(stat='identity', show.legend = TRUE, width = .8) +
  labs(fill = 'CA Beer Styles') +
ggtitle("Top California Beer Styles")
  
ggplotly(p)
  
```
Nearest Neighbor (KNN) Prediction for Ale

```{r}
#Filter data for IPA and Ale 
Aledata = full_data %>% mutate(IPA = grepl('IPA',full_data$Style), Ale  = grepl('Ale',full_data$Style)) %>% filter(IPA | Ale) %>% select(Ale,IPA,ABV,IBU, Style)

Aledata = Aledata %>% mutate(Alcohol_by_Volume = scale(ABV),biterness_scale = scale(IBU)) #Get Ranges of ABV and IBU data

summary(Aledata)

library(tidyverse)
library(caret)
library(class)

train = Aledata %>% select(Alcohol_by_Volume,biterness_scale,Ale,ABV,IBU) %>% na.omit()

dim(train) #944 rows

head(train)

trainIndex = sample(seq(1,944,1),100)

train2 = train[trainIndex,]

test2 = train[-trainIndex,]

AleType = test2$Ale

#IBU and ABV
classifications = knn(train2[,c(4,5)],test2[,c(4,5)],cl = train2$Ale,k = 28)
CM = table(classifications,AleType)
CM2 = confusionMatrix(CM)
CM2
CM2$overall[1] # accuracy
```

# Ale Classification by Alcohol Content and Bitterness
#Testing for Style of Ale based upon Alcohol Content and Bitterness
```{r, echo=FALSE, message = FALSE, warning=FALSE}
#Plot Classification test for Ales based upon ABV and IBU
p = Aledata %>% ggplot(mapping = aes(x=Alcohol_by_Volume,y=biterness_scale,color=Ale)) + geom_point(position = 'jitter') +
scale_color_manual(values=c("chartreuse4", "deepskyblue4")) +
ggtitle("Ale Classification - Alcohol Content and Bitterness")
ggplotly(p)
```

# Popular Styles by State
```{r ,echo=FALSE, message = FALSE, warning=FALSE}
#Filter Most Popluar Styles by State
pop_styles = full_data %>% select(Style) %>%
  count(Style) %>% slice_max(n,n = 10) %>% arrange(desc(n))

head(pop_styles)


p = pop_styles %>% 
  arrange(desc(n)) %>% 
  ggplot(mapping = aes(y=Style,x=n, fill = Style)) + geom_bar(stat='identity', show.legend = TRUE, width = .9) +
  labs(fill = 'Beer Styles') +
ggtitle("Top 10 Beer Styles in the US")
  
ggplotly(p)
```

Cities with Highest Number of Breweries
```{r}
#state_data -- Breweries Filtered by State - joined with FIPS code by State
#mutate - Join FIPS data with Breweries data
#trimws - remove Leading/Trailing Whitespace from column(State)
#fips = State by Abbreviation - ie. TX

city_data = breweries %>% group_by(City) %>% count(City) %>% arrange(desc(n)) 

head(city_data)
```



